Follow-up Samples

### 2017-01-16 submission/manuscript_outline_20161024.Rmd
streamline intro

Colorectal cancer (CRC) continues to be a leading cause of cancer related deaths and is the second most common cancer death among men aged 40-79 years of age [@jemal_cancer_2010; @haggar_colorectal_2009].  Over the last few years death due to the disease has seen a significant decrease, thanks mainly to improvements in screening [@jemal_cancer_2010].  However, despite this improvement there are still approximately 50,000 deaths from the disease a year [@haggar_colorectal_2009].  It is estimated that around 5-10% of all CRCs can be explained by autosomal dominant inheritance [@green_very_2007]. The vast majority of CRCs are not inherited and the exact etiology of the disease has not been well worked out [@haggar_colorectal_2009].  Although many risk factors have been identified [@haggar_colorectal_2009] and non-invasive screening techniques have started to be put into consistent use [@liao_application_2013; @johnson_multi-target_2016] there has been an consistent additional increase in the incidence of CRC in the younger population.

This increased incidence of CRC in the younger population is concerning since having either an adenoma or carcinoma increases ones risk for future adenomas or carcinomas [@laiyemo_short-_2013; @matsuda_five-year_2009; @ren_long-term_2016].  This increased risk can also carry with it an increased chance of mortality due to this recurrence [@loberg_long-term_2014; @freeman_natural_2013].  Therefore there has been a great amount of interest in early risk stratification tools [@lee_identification_2016; @richards_evidence-based_2016] that can help identify those they may be at most susceptibility to reccurence.  Concurrently, there has also been a lot of interest in new areas that could have a role in disease pathogenesis, such as the gut bacterial microbiome.

@@

In this study we investigated what happened to the bacterial microbiome before and after surgery for both adenoma and carcinoma individuals.  Our anlaysis includes both alpha and beta diversity analysis along with investigation of individual operational taxonomic units (OTUs).  We also utilized Random Forest models and observed how these models well as specific OTUs within this model performed on before (initial) and after (follow up) surgery samples.  We used these models to look for similar important OTUs to identify the most important ones for not only classifying initial and follow up samples but also lesion or normal.

>>>>

Colorectal cancer (CRC) continues to be a leading cause of cancer related deaths and is the second most common cancer death among men aged 40-79 years of age [@jemal_cancer_2010; @haggar_colorectal_2009].  Over the last few years death due to the disease has seen a significant decrease, thanks mainly to improvements in screening [@jemal_cancer_2010].  However, despite this improvement there are still approximately 50,000 deaths from the disease a year [@haggar_colorectal_2009].  It is estimated that around 5-10% of all CRCs can be explained by autosomal dominant inheritance [@green_very_2007]. The vast majority of CRCs are not inherited and the exact etiology of the disease has not been well worked out [@haggar_colorectal_2009].  Although many risk factors have been identified [@haggar_colorectal_2009] and non-invasive screening techniques have started to be put into consistent use [@liao_application_2013; @johnson_multi-target_2016] there has been an additional increase in the incidence of CRC in the younger population.

This increased incidence of CRC in the younger population is concerning since having either an adenoma or carcinoma increases ones risk for future adenomas or carcinomas [@laiyemo_short-_2013; @matsuda_five-year_2009; @ren_long-term_2016].  This increased risk can also carry with it an increased chance of mortality due to this recurrence [@loberg_long-term_2014; @freeman_natural_2013].  Therefore, there has been a great amount of interest in early risk stratification tools [@lee_identification_2016; @richards_evidence-based_2016] that can help identify those that may be at most susceptibility to reccurence.  Concurrently, there has also been a lot of interest in new areas that could have a role in disease pathogenesis, such as the gut bacterial microbiome.

@@

In this study we investigated what happened to the bacterial microbiome before and after surgery for indivdiuals with either adenoma and carcinoma.  Our anlaysis includes both alpha and beta diversity analysis along with investigation of individual operational taxonomic units (OTUs).  We also utilized Random Forest models and observed how these models as well as specific OTUs within this model performed on before (initial) and after (follow up) surgery samples.  We used these models to look for similar important OTUs to identify the crucial OTUs for not only classifying initial and follow up samples but also lesion or normal.

### 2017-01-16 submission/manuscript_outline_20161024.tex
streamline intro

Colorectal cancer (CRC) continues to be a leading cause of cancer
related deaths and is the second most common cancer death among men aged
40-79 years of age {[}1,2{]}. Over the last few years death due to the
disease has seen a significant decrease, thanks mainly to improvements
in screening {[}1{]}. However, despite this improvement there are still
approximately 50,000 deaths from the disease a year {[}2{]}. It is
estimated that around 5-10\% of all CRCs can be explained by autosomal
dominant inheritance {[}3{]}. The vast majority of CRCs are not
inherited and the exact etiology of the disease has not been well worked
out {[}2{]}. Although many risk factors have been identified {[}2{]} and
non-invasive screening techniques have started to be put into consistent
use {[}4,5{]} there has been an consistent additional increase in the
incidence of CRC in the younger population.

This increased incidence of CRC in the younger population is concerning
since having either an adenoma or carcinoma increases ones risk for
future adenomas or carcinomas {[}6--8{]}. This increased risk can also
carry with it an increased chance of mortality due to this recurrence
{[}9,10{]}. Therefore there has been a great amount of interest in early
risk stratification tools {[}11,12{]} that can help identify those they
may be at most susceptibility to reccurence. Concurrently, there has
also been a lot of interest in new areas that could have a role in
disease pathogenesis, such as the gut bacterial microbiome.

>>>>

Colorectal cancer (CRC) continues to be a leading cause of cancer
related deaths and is the second most common cancer death among men aged
40-79 years of age {[}1,2{]}. Over the last few years death due to the
disease has seen a significant decrease, thanks mainly to improvements
in screening {[}1{]}. However, despite this improvement there are still
approximately 50,000 deaths from the disease a year {[}2{]}. It is
estimated that around 5-10\% of all CRCs can be explained by autosomal
dominant inheritance {[}3{]}. The vast majority of CRCs are not
inherited and the exact etiology of the disease has not been well worked
out {[}2{]}. Although many risk factors have been identified {[}2{]} and
non-invasive screening techniques have started to be put into consistent
use {[}4,5{]} there has been an additional increase in the incidence of
CRC in the younger population.

This increased incidence of CRC in the younger population is concerning
since having either an adenoma or carcinoma increases ones risk for
future adenomas or carcinomas {[}6--8{]}. This increased risk can also
carry with it an increased chance of mortality due to this recurrence
{[}9,10{]}. Therefore, there has been a great amount of interest in
early risk stratification tools {[}11,12{]} that can help identify those
that may be at most susceptibility to reccurence. Concurrently, there
has also been a lot of interest in new areas that could have a role in
disease pathogenesis, such as the gut bacterial microbiome.
